---
title: "maximum.likelihood"
author: "Abby Robinson"
date: "2/19/2022"
output: html_document
---

Maximum Likelihood Function 
```{r}
##i = attacks on facsimile of a specific phenotype 
##jun = attacks on junonia 
likelihood = function(i, jun, prob){
  
  (i * log(prob / (prob+1)) ) + (jun * log(1/ (prob+1)) ) 

}

```

```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/quabbin.count.csv")
data <- read.csv(c, header = TRUE, sep = ",", stringsAsFactors = FALSE)
data
```
```{r}
sim <- subset(data, phase == "simultaneous", drop = FALSE)
ba <- subset(sim, species == "battus", drop = FALSE)
li <- subset(sim, species == "limenitis", drop = FALSE)
ju <- subset(sim, species == "junonia", drop = FALSE)

bl <- subset(sim, species %in% c("battus", "limenitis"), drop = FALSE)
jl <- subset(sim, species %in% c("junonia", "limenitis"), drop = FALSE)
bj <- subset(sim, species %in% c("battus", "junonia"), drop = FALSE)
```

```{r}

glm4 <- glm(data = sim, attacks ~ species, family = "poisson")
summary(glm4)

glm.bl <- glm(data = bl, attacks ~ species, family = "poisson")
summary(glm.bl)

glm.jl <- glm(data = jl, attacks ~ species, family = "poisson")
summary(glm.jl)

glm.bj <- glm(data = bj, attacks ~ species, family = "poisson")
summary(glm.bj)
```

```{r}
learning <- subset(data, phase == "learning", drop = FALSE)
testing <- subset(data, phase == "testing", drop = FALSE)
simultaneous <- subset(data, phase == "simultaneous", drop = FALSE)
learning.pld <- subset(learning, day %in% c(2,3,4), drop = FALSE)
testing.pld <- subset(testing, day %in% c(2,3,4), drop = FALSE)
```

```{r}
b <- subset(learning.pld, species == "battus", drop = FALSE)
bj <- subset(learning.pld, species == "junonia", drop = FALSE)

l <- subset(testing.pld, species == "limenitis", drop = FALSE)
lj <- subset(testing.pld, species == "junonia", drop = FALSE)
```

likelihood for limenitis totals 
```{r}
l.sum <- sum(l$attacks)
lim <- l$attacks

lj.sum <- sum(lj$attacks)
l.jun <- lj$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob.lj <- l.sum / (l.sum + lj.sum)
prob.lj

##likelihood scores for limenitis at each site for all four days
lim.likelihood <- likelihood(lim, l.jun, prob.lj)
li.like.score <- sum(lim.likelihood)
li.like.score
```

Likelihood for battus totals 
```{r}
b.sum <- sum(b$attacks)
bat <- b$attacks

bj.sum <- sum(bj$attacks)
b.jun <- bj$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking bat vs. jun: 

prob.bj <- b.sum / (b.sum + bj.sum)
prob.bj

##likelihood scores for battus at each site for all four days
bat.likelihood <- likelihood(bat, b.jun, prob.bj)
ba.like.score <- sum(bat.likelihood)
ba.like.score
```
###My problem: the proportion of attack rates calculated for maximum likelihood does not translate into significance 

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(li.like.score - ba.like.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 

##p-value = 6.286951e-06
##bonferroni correction? 
```

```{r}
pairwise.t.test(learning$attacks, learning$species, p.adjust.method="bonferroni")
```


****This analysis will show if there are any significant differences in day-to-day attack rates between the two species 
l1 v b1
```{r}
l.one <- subset(l, day == "1", drop = FALSE)
sum.l.one <- sum(l.one$attacks)
l.one.attack <- l.one$attacks

lj.one <- subset(lj, day == "1", drop = FALSE)
sum.lj.one <- sum(lj.one$attacks)
lj.one.attack <- lj.one$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob.lj.one <- sum.l.one / (sum.l.one + sum.lj.one)
prob.lj.one

##likelihood scores for limenitis at each site on day one 
lim.one.likelihood <- likelihood(l.one.attack, lj.one.attack, prob.lj.one)
li.one <- sum(lim.one.likelihood)
li.one
```

```{r}
b.one <- subset(b, day == "1", drop = FALSE)
sum.b.one <- sum(b.one$attacks)
b.one.attack <- b.one$attacks

bj.one <- subset(bj, day == "1", drop = FALSE)
sum.bj.one <- sum(bj.one$attacks)
bj.one.attack <- bj.one$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking bat vs. jun: 

prob.bj.one <- sum.b.one / (sum.b.one + sum.bj.one)
prob.bj.one

##likelihood scores for battus at each site on day one
bat.one.likelihood <- likelihood(b.one.attack, bj.one.attack, prob.bj.one)
ba.one <- sum(bat.one.likelihood)
ba.one
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(li.one - ba.one)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE )
p.value 

##p-value = 3.319921e-06
##bonferroni correction? 
##  I think the difference in sample sizes here is maybe messing uo the likelihood scores. 
```

####probability of birds attacking limenitis relative to junonia at transects 1, 2, 3, and the simultaneous experiment 

```{r}
l1 <- subset(l, transect == "1", drop = FALSE)
lj1 <- subset(lj, transect == "1", drop = FALSE)

l2 <- subset(l, transect == "2", drop = FALSE)
lj2 <- subset(lj, transect == "2", drop = FALSE)

l3 <- subset(l, transect == "3", drop = FALSE)
lj3 <- subset(lj, transect == "3", drop = FALSE)

sl <- subset(simultaneous, species == "limenitis", drop = FALSE)
sj <- subset(simultaneous, species == "junonia", drop = FALSE)
```

```{r}
l1.sum <- sum(l1$attacks)
lim1 <- l1$attacks

lj1.sum <- sum(lj1$attacks)
jun1 <- lj1$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob1 <- l1.sum / (l1.sum + lj1.sum)
prob1

##likelihood scores for limenitis at each site for all four days
l1.likelihood <- likelihood(lim1, jun1, prob1)
l1.score <- sum(l1.likelihood )
l1.score
```

```{r}
l2.sum <- sum(l2$attacks)
lim2 <- l2$attacks

lj2.sum <- sum(lj2$attacks)
jun2 <- lj2$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob2 <- l2.sum / (l2.sum + lj2.sum)
prob2

##likelihood scores for limenitis at each site for all four days
l2.likelihood <- likelihood(lim2, jun2, prob2)
l2.score <- sum(l2.likelihood )
l2.score
```

```{r}
l3.sum <- sum(l3$attacks)
lim3 <- l3$attacks

lj3.sum <- sum(lj3$attacks)
jun3 <- lj3$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob3 <- l3.sum / (l3.sum + lj3.sum)
prob3

##likelihood scores for limenitis at each site for all four days
l3.likelihood <- likelihood(lim3, jun3, prob3)
l3.score <- sum(l3.likelihood )
l3.score
```

```{r}
sl.sum <- sum(sl$attacks)
sim.lim <- sl$attacks

sj.sum <- sum(sj$attacks)
sim.jun <- sj$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking lim vs. jun: 

prob4 <- sl.sum / (sl.sum + sj.sum)
prob4

##likelihood scores for limenitis at each site for all four days
l4.likelihood <- likelihood(sim.lim, sim.jun, prob4)
l4.score <- sum(l4.likelihood )
l4.score ###The low attack rate on limenitis is getting obscured by the high attack rate on junonia in this score 
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l1.score - l2.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l1.score - l4.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 

```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l2.score - l3.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l1.score - l3.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l2.score - l4.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(l3.score - l4.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```

Compare: 
p-values without bonferroni correction: 

l1 v l2  0.07987582
l1 v l3  0.008397561
l2 v l3  0.04888671

l1 v l4 0.00401174
l2 v l4 0.02244356
l3 v l4 0.2484803

attack rates on limenitis are not significantly different between transect 1 and transect 2, and between transect 3 and transect 4, but transects 1&2 are significantly different from transects 3&4
###transect 3 might just be a little weird because it has a lower attack rate overall....

####should I get a probability for all limenitis v. all junonia and compare individual transects to that probability? 

```{r}
mimic <- subset(data, phase %in% c("testing", "simultaneous"), drop = FALSE)

limenitis <- subset(mimic, species == "limenitis", drop = FALSE)
l <-sum(limenitis$attacks)

junonia <- subset(mimic, species == "junonia", drop = FALSE)
j <-sum(junonia$attacks)

prob <- l / (l + j)
prob
```

##I think the fact that our junonia counts vary for each comparison is fucking things up... 

The probabilities make sense but the likelihood scores don't... it seems like the attack total at each comparison is too important, and attacks on limenitis are being scored basically the same as attacks on junonia, which doesn't make sense.... 

Something that has a higher number of attacks overall is going to have a higher likelihood score

Likelihood for battus totals 
```{r}
b4 <- subset(simultaneous, species == "battus", drop = FALSE)
j4 <- subset(simultaneous, species == "junonia", drop = FALSE)

b4.sum <- sum(b4$attacks)
bat4 <- b4$attacks

j4.sum <- sum(j4$attacks)
jun4 <- j4$attacks

###maximum likelihood calculated for each species 
#for the probability of attacking bat vs. jun: 

prob.bj4 <- b4.sum / (b4.sum + j4.sum)
prob.bj4

##likelihood scores for battus at each site for all four days
bat4.likelihood <- likelihood(bat4, jun4, prob.bj4)
b4.score <- sum(bat4.likelihood)
b4.score
```

```{r}
##This is the likelihood ratio test to get significance level 
DIFF <- abs(b4.score - l4.score)
DIFF 

G <- DIFF*2
G 

p.value <- pchisq(G, 1, lower.tail = FALSE)
p.value 
```
