---
title: "PAVO"
author: "Abby Robinson"
date: "10/19/2022"
output: html_document
---

```{r}
library(curl)
library(pavo)
```

###Making the data frame 
```{r}
p <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/spectral_reflectance_data.csv")
specs <- read.csv(p, header = TRUE, sep = ",")
specs <- subset(specs, wavelength >= 300 & wavelength <= 701, drop = FALSE)
lim <- subset(specs, species == "limenitis", drop = FALSE) 
bat <- subset(specs, species == "battus", drop = FALSE) 

lim_real <- subset(lim, specimen.type == "real.wing", drop = FALSE) 
  lim_real_1 <- subset(lim_real, sample.number == "2", drop = FALSE) 

lim_fac <- subset(lim, specimen.type == "facsimiles", drop = FALSE) 
  lim_fac_1 <- subset(lim_fac, sample.number == "1", drop = FALSE) 
  lim_fac_2 <- subset(lim_fac, sample.number == "2", drop = FALSE) 
  lim_fac_3 <- subset(lim_fac, sample.number == "3", drop = FALSE) 

bat_real <- subset(bat, specimen.type == "real.wing", drop = FALSE) 
    bat_real_1 <- subset(bat_real, sample.number == "1", drop = FALSE) 
    
bat_fac <- subset(bat, specimen.type == "facsimiles", drop = FALSE) 
  bat_fac_1 <- subset(bat_fac, sample.number == "4", drop = FALSE) 
  bat_fac_2 <- subset(bat_fac, sample.number == "5", drop = FALSE) 
  bat_fac_3 <- subset(bat_fac, sample.number == "6", drop = FALSE) 
  
###################
  
lr <- subset(lim_real_1, select = c(wavelength, reflectance) )

lf1 <- subset(lim_fac_1, select = c(wavelength, reflectance) )
lf2 <- subset(lim_fac_2, select = c(wavelength, reflectance) )
lf3 <- subset(lim_fac_3, select = c(wavelength, reflectance) )

br <- subset(bat_real_1, select = c(wavelength, reflectance) )

bf1 <- subset(bat_fac_1, select = c(wavelength, reflectance) )
bf2 <- subset(bat_fac_2, select = c(wavelength, reflectance) )
bf3 <- subset(bat_fac_3, select = c(wavelength, reflectance) )
```


```{r}

# Convert to rspec object
limreal <- as.rspec(lr)
batreal <- as.rspec(br)
###designed the facsimiles based on spectral reflectance from these individuals 

limreal <- procspec(
  limreal,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)

batreal <- procspec(
  batreal,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)

limfac.1 <- as.rspec(lf1)
limfac.2 <- as.rspec(lf2)
limfac.3 <- as.rspec(lf3)

batfac.1 <- as.rspec(bf1)
batfac.2 <- as.rspec(bf2)
batfac.3 <- as.rspec(bf3)

summary(limreal, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(batreal, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(limfac.1, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(limfac.2, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(limfac.3, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(batfac.1, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(batfac.2, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
summary(batfac.3, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
```

```{r}
wl <- limreal$wl
lim.real <- limreal$reflectance
bat.real <- batreal$reflectance

lim.fac.1 <- limfac.1$reflectance
lim.fac.2 <- limfac.2$reflectance
lim.fac.3 <- limfac.3$reflectance

bat.fac.1 <- batfac.1$reflectance
bat.fac.2 <- batfac.2$reflectance
bat.fac.3 <- batfac.3$reflectance

lim.fac <- cbind(wl, lim.fac.1, lim.fac.2, lim.fac.3)
bat.fac <- cbind(wl, bat.fac.1, bat.fac.2, bat.fac.3)

lim.fac <- as.rspec(lim.fac)
bat.fac <- as.rspec(bat.fac)
```


```{r}
mean.lim.fac <- aggspec(lim.fac, FUN = mean)
mean.bat.fac <- aggspec(bat.fac, FUN = mean)

lim.fac <- procspec(
  mean.lim.fac,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)

bat.fac <- procspec(
  mean.bat.fac,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)

lim.facsimile <- lim.fac$dat
bat.facsimile <- bat.fac$dat

specs <- cbind(wl, lim.real, bat.real, lim.facsimile, bat.facsimile )

is.rspec(specs)
specs <- as.rspec(specs)
is.rspec(specs)
specs

summary(specs, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
```

```{r}
par(mfrow = c(2, 2))
peakshape(specs, plot = TRUE, lim = c(350, 600))
```


```{r}
bluetit_sense <- sensdata(visual = "bluetit", achromatic = "bt.dc")
head(bluetit_sense)
tail(bluetit_sense)

uv_sense <- sensdata(visual = "avg.uv", achromatic = "bt.dc")
head(uv_sense)
tail(uv_sense)

v_sense <- sensdata(visual = "avg.v", achromatic = "bt.dc")
head(v_sense)
tail(v_sense)
```


```{r}
vismodel <- vismodel(specs,
  visual = "avg.v", achromatic = "ch.dc",
  illum = "forestshade", relative = TRUE
)
vismodel.cd <- coldist(vismodel)
vismodel.cd
```

Colour distances are obtained by weighting the Euclidean distance of the photoreceptor quantum catches by the Weber fraction of the cones (ΔS). These measurements are in units of Just Noticeable Differences (JNDs), where distances over a certain threshold (usually 1) are considered to be discernible under the conditions considered (e.g., backgrounds, illumination). 

###Where dS is the chromatic contrast (ΔS) and dL is the achromatic contrast (ΔL). 

```{r}
specsmod.cd <- coldist(specsmod)
```
```{r}
# Convert points to Cartesian coordinates in which Euclidean distances are 
# noise-weighted.
propxyz <- jnd2xyz(specsmod.cd)

# Plot the floral spectra in 'noise-corrected' space
plot(propxyz)
```


```{r}
a <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/battus_real_wing_spectra.csv")
bat.real <- read.csv(a, header = TRUE, sep = ",")
bat.real <- subset(bat.real, wavelength >= 300 & wavelength <= 701, drop = FALSE)
bat.real <- as.rspec(bat.real)
bat.real <- procspec(
  bat.real,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)
```

```{r}
b <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/battus_facsimile_spectra.csv")
bat.fac <- read.csv(b, header = TRUE, sep = ",")
bat.fac <- subset(bat.fac, wavelength >= 300 & wavelength <= 701, drop = FALSE)
bat.fac <- as.rspec(bat.fac)
bat.fac <- procspec(
  bat.fac,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)
```

```{r}
c <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/limenitis_real_wing_spectra.csv")
lim.real <- read.csv(c, header = TRUE, sep = ",")
lim.real <- subset(lim.real, wavelength >= 300 & wavelength <= 701, drop = FALSE)
lim.real <- as.rspec(lim.real)
lim.real <- procspec(
  lim.real,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)
```

```{r}
d <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/limenitis_facsimile_spectra.csv")
lim.fac <- read.csv(d, header = TRUE, sep = ",")
lim.fac <- subset(lim.fac, wavelength >= 300 & wavelength <= 701, drop = FALSE)
lim.fac <- as.rspec(lim.fac)
lim.fac <- procspec(
  lim.fac,
  opt =  "smooth",
  fixneg = "zero",
  span = 0.25,
  bins = 20
)
```

```{r}
wl <- bat.real$wl
bat.real <- bat.real$reflectance
lim.real <- lim.real$reflectance
bat.fac <- bat.fac$reflectance
lim.fac <- lim.fac$reflectance

specs <- cbind(wl, bat.real, lim.real, bat.fac, lim.fac)
specs <- as.rspec(specs)

summary(specs, subset = c('B1', 'B2', 'B3', 'H1', 'H2', 'H3'))
```
```{r}
par(mfrow = c(2, 2))
peakshape(specs, plot = TRUE, lim = c(350, 700))
```
```{r}
vismodel <- vismodel(specs,
  visual = "avg.v", achromatic = "ch.dc",
  illum = "D65", relative = TRUE
)
vismodel.cd <- coldist(vismodel)
vismodel.cd
```
```


