---
title: "Temporal Dynamics of Predator Learning in a Batesian Mimicry Complex"
author: "Abby Robinson"
date: "12/16/2022"
output: html_document
---

## [1] Did avian predators learn to avoid facsimiles of Battus philenor during 4-day experiments, compared to the undefended facsimiles of the control? {.tabset}

In order to address how the time delay will affect attack rates on the mimics, we first need to show that our training experiment was effective and birds learned to avoid facsimilies of battus philenor. 

I hypothesize that birds will learn to avoid the chemically defended models faster than the control, and that we will therefore see lower attack rates on the model in the last three days of the experiment compared to the control. 

Because this initial question of the project is focused on the training phase of the experiment, each of the three time-delayed experiments (four, two, one) are essentially identical replicates of each other, and therefore this data can be pooled to increase power 

Packages
```{r}
library(curl)
library(ggplot2)
library(lme4)
library(DHARMa)
library(emmeans)
library("glmmTMB")
library(patchwork)
```

```{r}
m <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/Quabbin_Model_Data.csv")
model <- read.csv(m, header = TRUE, sep = ",")
head(model)
```

```{r}
# make field.day variable factor
class(model$field.day) 
model$field.day <- as.factor(model$field.day)
class(model$field.day)

# make site variable factor 
class(model$field.site)
model$field.site <- as.factor(model$field.site)
class(model$field.site)

# make sure attack variable is integer 
class(model$attacks) 

# make species variable factor 
class(model$species)
model$species <- as.factor(model$species)
class(model$species)

# make treatment variable factor 
class(model$treatment)
model$treatment <- as.factor(model$treatment)
class(model$treatment)
```

# make experimental.day variable factor 
class(model$experiment.day) 
count$experiment.day <- as.factor(count$experiment.day)
class(count$experiment.day)

```{r}
# histogram of attacks 
hist(count$attacks)  #concerns about zero inflation based on histogram? 

# check variance and mean to assess for over /  under - dispersion 
var(count$attacks)
mean(count$attacks) # variance is greater than mean, which indicates that over-dispersion might be an issue 

```

```{r}
mod5 <- glmmTMB(attacks ~ experiment.day*species + treatment*species + (1|field.site/field.day), data=count,ziformula=~1,family=poisson) 
summary(mod5)
check_singularity(mod5) 

# no singularity issues 
# no convergence issues 
```

```{r}
mod5simulation <- simulateResiduals(fittedModel = mod5, plot = T)

testUniformity(mod5simulation) 
testOutliers(mod5simulation)
testDispersion(mod5simulation) 
testZeroInflation(mod5simulation) 
```
```{r}
## MODEL 5
post.hoc1 <- emmeans(mod5, pairwise ~ experiment.day|species) 

# do I still need to do a bonferroni correction if I am no longer running "extra" pairwise comparisons 
post.hoc1$contrasts %>%
     rbind() # bonferroni adjustment for multiple comparisons 
# bonferroni didn't change significance 
```
```{r}
m <- curl("https://raw.githubusercontent.com/butterfliesrcool/Quabbin_Data/main/Quabbin_Mimic_Data.csv")
mim.data <- read.csv(m, header = TRUE, sep = ",")
head(mim.data)
# count data looses some information that binomial data would provide, but binomial data would likely not work here due to the low overall attack rates 
# survival analysis likely didn't work with this data because attack rates are too low for this data to be analyzed as binomial data 
```

```{r}
# make experimental.day variable factor 
class(mim.data$experiment.day) 
mim.data$experiment.day <- as.factor(mim.data$experiment.day)
class(mim.data$experiment.day)

# make field.day variable factor
class(mim.data$field.day) 
mim.data$field.day <- as.factor(mim.data$field.day)
class(mim.data$field.day)

# make site variable factor 
class(mim.data$field.site)
mim.data$field.site <- as.factor(mim.data$field.site)
class(mim.data$field.site)

# make sure attack variable is integer 
class(mim.data$attacks) # only thing that should be an integer 

# make species variable factor 
class(mim.data$species)
mim.data$species <- as.factor(mim.data$species)
class(mim.data$species)

# make time.delay variable factor 
class(mim.data$treatment)
mim.data$treatment <- as.factor(mim.data$treatment)
class(mim.data$treatment)
```

```{r}
mod2 <- glmmTMB(attacks ~ species*treatment + (1|field.site) + (1|field.day), data=mim.data,ziformula=~1,family=poisson) 
summary(mod2)
check_singularity(mod2) 
```
```{r}
mod3 <- glmmTMB(attacks ~ species*treatment + species*experiment.day + (1|field.site) + (1|field.day), data=mim.data,ziformula=~1,family=poisson) 
summary(mod3)
check_singularity(mod3) 
```
```{r}
mod3simulation <- simulateResiduals(fittedModel = mod3, plot = T)
testUniformity(mod3simulation) 
testOutliers(mod3simulation)
testDispersion(mod3simulation) 
testZeroInflation(mod3simulation) 
```
```{r}
post.hoc2 <- emmeans(mod3, pairwise ~ species|treatment)
post.hoc3 <- emmeans(mod3, pairwise ~ experiment.day|species)
```


